Molecular dynamics simulation of synchronization in driven particles

### 2021-02-26 manuscript/manuscript_ajp_2020.tex
shifted phi by phi0 to be centered about zero

  To study synchronization patterns, 
  it is useful to compare
  mode-locked quantities %periodic signals
  in a two dimensional phase plot. 
  In a non-driven system, an appropriate
  phase space is particle velocity $v_y$ versus position $y$.  
  For a particle driven parallel to the $y-$direction,
  %the particle velocity and position increase
  %in time. % due to $F^{dc}$.
  %Since a system does not form
  %a closed loop in
  %$v_y$ vs $y$ space,
  we define phase variables 
  to account for the net increase in velocity
  and position.
  The phase position is
  defined as 
  %within the substrate period as 
  %We define a phase variable 
  \begin{equation}
    %(y mod \lambda) / \lambda
    \phi(t) = (y(t)-\langle v_y \rangle t)/\lambda
  \end{equation}
  centered about the average particle displacement $\langle v_y \rangle t$
  and normalized by the substrate period $\lambda$ \cite{Juniper2015}.
  For clarity we have dropped a prefactor of $2\pi$ 
  to compare $\phi(t)$ to $\lambda$.
  %In phase diagrams we plot $\phi(t)$ %phase
  %versus
  The phase velocity is
  \begin{equation}
    %(y mod \lambda) / \lambda
    \dot{\phi}(t) =(v_y(t)-\langle v_y \rangle) /\lambda.  % 2\pi
  \end{equation}
  With no landscape force, 
  the phase velocity will be zero when $F^d(t) = F^{dc}$.
  %Juniper %REF 53
  %and easily generated with a computer code 
  %or oscilloscope \cite{Tong1997}.
  When velocity
  is mode-locked to %repeat periodically
  a spatial location
  a closed loop appears in  
  phase space. % showing the relationship
  %between two periodic functions
  The pattern is determined by the mode,
  or frequency ratio,
  so a 1:1 mode appears as a circle or oval. 
  Nodes appear %in the oval 
  for higher modes,
  sometimes forming figure-eights
  or other recognizable patterns.
  A system that is nearly phase locked
  will appear as an unclosed loop.
  Such quasiperiodic systems are
  not fully synchronized
  so the position-velocity relationship
  shifts in time.
  In Fig.~\ref{fig:phase}, 
  all cases exhibit some line thickening, 
  indicating the particle location and velocity
  are not fully synchronized,
  although this can be tuned by
  calculating $\langle v_y \rangle$ over
  shorter time ranges.
  %and the shifts in phase space represent
  %small changes in system behavior.
  %With a long simulation time,
  %a quasiperiodic phase diagram
  %fills the plotting region due to the gradual shift
  %in the velocity-position relationship.
  %with a frequency generator,
  %where two combined frequencies at perpendicular orientations
  %\footnote{The term phase
  %has two scientific meanings:  
  %oscillation phase $\phi_0$ is the constant of the argument
  %in a periodic function
  %$\sin{(\omega t + \phi_0)}$ and phase space is
  %used to characterize oscillating systems with
  %parametric plots.}.

@@

  In Fig.~\ref{fig:shapiro}
  we sweep the constant driving force $F^{dc}$
  for a fixed amplitude $F^{ac}$ and frequency.
  Fig.~\ref{fig:shapiro}(a) shows $F^{ac}= 0.0$ and (b) shows $F^{ac}=0.4$
  with frequency $f=0.01$.

  The particle does not phase lock since there is no periodicity to $F^d(t)$,
  but does exhibit undulations in $\langle v_y \rangle$
  due to the corregations in the substrate.
  %WAIT, THAT DOESN'T MAKE SENSE!
  %WHAT IS WRONG WITH YOUR MD SIMULATION?
  When $F^{ac}=0.4$, 
  mode-locked steps appear in $\langle v_y \rangle$ vs. $F^{dc}$.
  At low values of $F^{dc}$ the particle may move backwards,
  distorting the steps, as seen below $F^{dc}<0.3$
  Between $0.3< F^{dc}<0.5$ the system exhibits phase locked steps.
    \begin{figure} %[h!]
      \centering
      \includegraphics[width=\columnwidth]{{fig7}.pdf}
      %\includegraphics[width=\columnwidth]{{fig7_sweep_FDC_vs_vx}.pdf}
      \caption{
        $\langle v_y \rangle$ vs. $F^{dc}$ of a particle with
        $F^{ac}=0.1$ (blue), $F^{ac}=0.2$ (orange),
        $F^{ac}=0.1$ (green) and $F^{ac}=0.2$ (red)
        with frequency $f=0.01$.
        The substrate is as in Fig.~\ref{fig:0}
        with $A_p = 0.1$.
      }
      \label{fig:shapiro}
    \end{figure}
  %In Fig.~\ref{fig:hops}
%\item Frenkel-Kontorova and competing length scales for low temperature systems where thermal fluctuations are negligible
%  \item The Fokker-Planck Equation
%  \item Aubrey-Kontorova + commensurability
%  \item nanotribology
 %   kink motion
 %   Vanossi et al J.Phys.: Condens. Matter 19 (2007)
  %\item
  %\subsection{Visualizing motion}
  %\label{ex:animation}
  %Drawing the contour 
  %The code for generating
  %a two dimensional colored plot
  %of the potential landscape
  %is calculated by evaluating
  %the analytic function in Eq.~\ref{eq:ysubstrate}
  %for a grid of values $(x_n,y_n)$.
  %is included in the supplementary material,
  %{\it I have python code that does all of this, but I don't have it
  %  set up to play nice with my new python MD code} 
  %Also the animation library in python...

>>>>

  To study synchronization patterns, 
  it is useful to compare
  mode-locked quantities %periodic signals
  in a two dimensional phase plot. 
  In a non-driven system, an appropriate
  phase space is particle velocity $v_y$ versus position $y$.  
  For a particle driven parallel to the $y-$direction,
  %the particle velocity and position increase
  %in time. % due to $F^{dc}$.
  %Since a system does not form
  %a closed loop in
  %$v_y$ vs $y$ space,
  we define phase variables 
  to account for the net increase in velocity
  and position.
  The phase position is
  defined as 
  %within the substrate period as 
  %We define a phase variable 
  \begin{equation}
    %(y mod \lambda) / \lambda
    \phi(t) = 2\pi (y(t)-\langle v_y \rangle t)/\lambda
  \end{equation}
  centered about the average particle displacement $\langle v_y \rangle t$
  and normalized by the substrate period $\lambda$ \cite{Juniper2015}.
  For clarity our plots are normalized by the prefactor $2\pi$. 
  %to compare $\phi(t)$ to $\lambda$.
  %In phase diagrams we plot $\phi(t)$ %phase
  %versus
  The phase velocity is
  \begin{equation}
    %(y mod \lambda) / \lambda
    \dot{\phi}(t) =2\pi (v_y(t)-\langle v_y \rangle) /\lambda.  % 
  \end{equation}
  With no landscape force, 
  the phase velocity will be zero when $F^d(t) = F^{dc}$.
  %Juniper %REF 53
  %and easily generated with a computer code 
  %or oscilloscope \cite{Tong1997}.
  When velocity
  is mode-locked to %repeat periodically
  a spatial location
  a closed loop appears in  
  phase space. % showing the relationship
  %between two periodic functions
  The pattern is determined by the mode,
  or frequency ratio,
  so a 1:1 mode appears as a circle or oval. 
  Nodes appear %in the oval 
  for higher modes,
  sometimes forming figure-eights
  or other recognizable patterns.
  A system that is nearly phase locked
  will appear as an unclosed loop.
  Such quasiperiodic systems are
  not fully synchronized
  so the position-velocity relationship
  shifts in time.
  In Fig.~\ref{fig:phase}, 
  all cases exhibit some line thickening, 
  indicating the particle location and velocity
  are not fully synchronized,
  although this can be tuned by
  calculating $\langle v_y \rangle$ over
  shorter time ranges.
  %and the shifts in phase space represent
  %small changes in system behavior.
  %With a long simulation time,
  %a quasiperiodic phase diagram
  %fills the plotting region due to the gradual shift
  %in the velocity-position relationship.
  %with a frequency generator,
  %where two combined frequencies at perpendicular orientations
  %\footnote{The term phase
  %has two scientific meanings:  
  %oscillation phase $\phi_0$ is the constant of the argument
  %in a periodic function
  %$\sin{(\omega t + \phi_0)}$ and phase space is
  %used to characterize oscillating systems with
  %parametric plots.}.

@@

  In Fig.~\ref{fig:shapiro}
  we sweep the constant driving force $F^{dc}$
  for a fixed amplitude $F^{ac}$ and frequency $f=0.1$.
  In the high frequency limit,
  the steps are separated
  by linearly increasing regions.

  Fig.~\ref{fig:shapiro}(a) shows $F^{ac}= 0.0$ and (b) shows $F^{ac}=0.4$
  with frequency $f=0.01$.

  The particle does not phase lock since there is no periodicity to $F^d(t)$,
  but does exhibit undulations in $\langle v_y \rangle$
  due to the corregations in the substrate.
  %WAIT, THAT DOESN'T MAKE SENSE!
  %WHAT IS WRONG WITH YOUR MD SIMULATION?
  When $F^{ac}=0.4$, 
  mode-locked steps appear in $\langle v_y \rangle$ vs. $F^{dc}$.
  At low values of $F^{dc}$ the particle may move backwards,
  distorting the steps, as seen below $F^{dc}<0.3$
  Between $0.3< F^{dc}<0.5$ the system exhibits phase locked steps.
    \begin{figure} %[h!]
      \centering
      \includegraphics[width=\columnwidth]{{fig7}.pdf}
      %\includegraphics[width=\columnwidth]{{fig7_sweep_FDC_vs_vx}.pdf}
      \caption{
        $\langle v_y \rangle$ vs. $F^{dc}$ of a particle with
        $F^{ac}=0.1$ (blue), $F^{ac}=0.2$ (orange),
        $F^{ac}=0.1$ (green) and $F^{ac}=0.2$ (red)
        with frequency $f=0.0$.
        The substrate is as in Fig.~\ref{fig:0}
        with $A_p = 0.1$.
      }
      \label{fig:shapiro}
    \end{figure}
  %In Fig.~\ref{fig:hops}
%\item Frenkel-Kontorova and competing length scales for low temperature systems where thermal fluctuations are negligible
%  \item The Fokker-Planck Equation
%  \item Aubrey-Kontorova + commensurability
%  \item nanotribology
 %   kink motion
 %   Vanossi et al J.Phys.: Condens. Matter 19 (2007)
  %\item
  %\subsection{Visualizing motion}
  %\label{ex:animation}
  %Drawing the contour 
  %The code for generating
  %a two dimensional colored plot
  %of the potential landscape
  %is calculated by evaluating
  %the analytic function in Eq.~\ref{eq:ysubstrate}
  %for a grid of values $(x_n,y_n)$.
  %is included in the supplementary material,
  %{\it I have python code that does all of this, but I don't have it
  %  set up to play nice with my new python MD code} 
  %Also the animation library in python...
